---
title: "Wikipedia Article Quality Prediction"
subtitle: "Final Report: DLSS"
author: 
  - name: "Nafiß Tavakol, Kuon Ito, Lorenz, Rückert, Marius Helten"
    affiliation: "1406810"
editor: visual
date: "`r format(Sys.Date(), '%B %d, %Y')`" 
format:
  pdf:
    fig-width: 6
    fig-height: 3
    keep-tex: false  
    colorlinks: true
    documentclass: report
    links-as-notes: true
    include-in-header: 
      text: |
        \usepackage{float}
header-includes:
  - \usepackage{titling}  
  - \pretitle{\begin{center}\LARGE\bfseries} 
  - \posttitle{\end{center}}  
  - \preauthor{\begin{center} \large} 
  - \postauthor{\end{center}} 
  - \predate{\begin{center}\large} 
  - \postdate{\begin{figure}[H]
      \centering
      \includegraphics[width=1.0\textwidth]{Images/placeholder.png}
    \end{figure}
    \end{center}} 
bibliography: references.bib
cite-method: citeproc
link-citations: true
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
data_descriptive <- read.csv("/Users/ramius/Desktop/CodeVault/01_Project/Uni/DLSS_DeepLearningforSocialScientists/Final_Project/DLSS-WAQP/2_FinalReport/Tables/Wiki_metrics.csv")

data_homophily <- read.csv("/Users/ramius/Desktop/CodeVault/01_Project/Uni/DLSS_DeepLearningforSocialScientists/Final_Project/DLSS-WAQP/2_FinalReport/Tables/Wiki_assortativity.csv")
```

# Introduction

(Problem description and approach)

Citing this way: [-@hanGraphNeuralNetworks2020], [-@montiFakeNewsDetection2019]

# Data Analysis

## Dataset Description

(Features and Target Variable, Article based)

## Wikipedia Network

### Graph Description

-   single conencted component and a directed network
-   sparse network
-   reasonably clustered
-   pretty sizeable sample but by no means exhaustive for all of wikipedia

```{r, echo=FALSE}
knitr::kable(data_descriptive, caption = "Network Descriptive Metrics", 
              digits = 5, 
              format.args = list(scientific = FALSE))
```

-   Degree Distribution
-   reasonably similar to powerlaw or lognormal - in any case heavy tailed (very few nodes with a lot of connections), pretty normal for many internet networks, rich-get richer effect

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth, height=5cm, keepaspectratio]{Images/Degree_PowerLaw.png}
    \caption{Wikipedia Graph - Degree Distribution and Power Law Analysis}
    \label{fig:pca_combined}
\end{figure}

-   for none of the tested attributes assortativity could be measured, which is very surprising and also pretty disheartneing
-   basically similar nodes are not neighbors which takes away one of the main theoretic assumptions

```{r, echo=FALSE}
knitr::kable(data_homophily, caption = "Network Assortativity", 
              digits = 4, 
              format.args = list(scientific = FALSE))
```

### Graph based Features

-   For one we used the conenctions in the network as part of the data that all the neural networks except for the MLP-Baseline were trained
-   

## Preprocessing

During the pre-processing we face two important challenges. The first was the highly imbalanced target variable with only a very small fraction of high-quality articles. The second challenge were the heavily skewed distributions, especially for the graph based features.

Regarding our target variable we tried switching from a classification to a regression problem. This allowed us to circumvent class counts by treating the ordinal categorical attributes as a numerical attribute. The heavily skewed distribution was log-transformed in order to obtain a less skewed distribution. After a brief evaluation, this approach proved to be flawed because the model just made average predictions. We concentrated on the classification approach. A first measure was to aggregated the categories into three ordinal classes from the initial count of nine. This improved class frequencies to a reasonable degree where training and predictions became possible.

The features had to be preprocessed since the numerical ranges were not uniform and fit for training. Here normalization could have solved the problem. However the distributions particularly for the network metrics are immensely positvely skewed. Different kinds such as standard, minmax, robust and robust-log scaling proved to be ineffective to generate reasonably spread distributions. The only approach that brought reasonable results was quantile scaling. Quantile scaling transforms data by mapping each value to its percentile rank, creating a uniform distribution where extreme outliers get compressed while preserving relative order.

# Methods

(Model architectures, training procedures, evaluation metrics)
## Model architectures

## GraphSAGE

GraphSAGE (Graph Sample and Aggregate) is an inductive graph neural network that learns node representations by aggregating information from sampled neighbors. In the presented implementations, three variants are used.

For classification, the model stacks three `SAGEConv` layers with batch normalization, ReLU activations, and dropout. The final layer outputs raw logits corresponding to class labels. For regression, the same structure is adopted, except that the last layer outputs a single value per node, producing predictions for continuous targets.

An improved regression variant introduces two changes: batch normalization is replaced with `LayerNorm` to stabilize training across varying graph structures, and the ReLU activation is substituted with GELU for smoother nonlinear transformations. These modifications aim to enhance convergence and predictive performance compared to the basic GraphSAGE regression model.

#### Residual GraphSAGE

ResidualGraphSAGE extends the standard GraphSAGE architecture with several improvements designed to enhance stability and expressiveness. First, an input projection layer is added so that raw features are linearly mapped into a common hidden space before message passing begins. Each SAGE layer is followed by layer normalization and GELU activation, which provide smoother and more stable transformations than batch normalization with ReLU. Residual connections are introduced around every SAGE layer, allowing gradients to flow more easily through the network and mitigating the problem of vanishing information in deeper models. Finally, an output projection maps the learned hidden representations into the desired output space. These modifications improve convergence and enable deeper architectures compared to the vanilla GraphSAGE model.


## Graph Convolutional Network (GCN)

The Graph Convolutional Network (GCN) updates node representations by aggregating information from neighbors through graph convolution operations. In its basic form, the model begins with an input projection that maps raw node features into a hidden space. Stacked GCN layers then propagate and transform these features across the graph, each followed by a non-linear activation such as ReLU and dropout for regularization. Finally, an output layer projects the hidden representation to the class space, producing logits for classification. This simple yet effective architecture serves as a baseline for many graph learning tasks.


## Graph Attention Network (GAT)

The Graph Attention Network extends the idea of graph convolutions by introducing attention mechanisms to weight the contribution of neighboring nodes. The model starts with an input projection into a hidden representation and applies multiple GAT layers, each computing attention coefficients over neighbors and aggregating their features accordingly. Intermediate layers use ELU activations, dropout for regularization, and layer normalization for stability. In the final layer, attention heads are averaged rather than concatenated, and the output is mapped directly to the target dimension. This architecture enables the model to focus on the most informative neighbors while maintaining flexibility for both classification and regression tasks.


# Results

(Performance comparison tables, learning curves, confusion matrices)

# Discussion and Conclusion

(Performance Comparison, Interpret the results in the context of social network theory, Key Findings and Implications)

# References

## Code and Data

-   API´s
-   Python Packages

## Literature

-   Citeable papers